---

title: "PSTAT 174 HW4"
author: "Kendall Brown"
date: "Winter 2018"
output: pdf_document
---

```{r}
acc.death=read.table("C:/Users/kebro/Desktop/Pstat 174/deaths .txt",header =F)
acc.death.ts=ts(acc.death)
plot.ts(acc.death.ts)
acf(acc.death.ts)
pacf(acc.death.ts)
mean(acc.death.ts)
var(acc.death.ts)
```
From the time series plot we see a constant mean with seasonal variance.
From the ACF we observe oscilation which implies an auto-regressive model.
From the PACF we observe spikes at lags 3,6,9, 11, and 13.

```{r}
acc.death.diff=diff(acc.death.ts,12)
plot.ts(acc.death.diff)
mean(acc.death.diff)
var(acc.death.diff)
acf(acc.death.diff)
pacf(acc.death.diff)
```
With mean -171 and variance 288714 we see the detrended and non-seasonal transform of the data.
From the ACF we observe oscilation with no decay implying an auto-regressive model.
From the PACF we observe a single spike at lag 2.
Differencing to remove seasonality was a good idea as it decreased observed variance by a more than significant amount. 
